Say ``Hello World!''.

### 2025-05-02 Manuscript/Manuscript.tex
Removing newpages
Draft++

% LANGUAGE NAME
\newcommand{\langName}{DroLang}

@@

% MAIN BODY
\section*{Background}
The simplest way to describe my project is a programming language which appears naturalistic. The original programming languages were punch-card based and tied in to the machinery they worked on. Since then, new languages have evolved from low level machine code into higher level abstracted systems. The evolution from Fortran to C to Python (to name a few) showcases how languages have moved in a direction allowing for programmers to more abstractly tell the computer what to do. \langName{} is the natural next step in that evolution.

@@

While a fully implemented natural programming language has the potential to solve many problems, the development of \langName{} pimarily focused on the problem of accessibility. Software developers are the first thought when it comes to programming, but many others need to use computers to accomplish complicated tasks in research and other areas. Many researchers use various programs to collect and analyze data, run simulations, and perform other tasks. However, for many, programming is not familiar or easy to grasp. In my own experience teaching programming to children at a summer camp, things as elementary as variables can be incredibly baffling when you've never been exposed to it.

@@

I have delved deep into parsers and semantic theory and learned how to make them work together. Having taken classes in syntax, semantics, and compilers, this is not an entirely new area for me, but my understanding of these topics have expanded beyond the scope of those classes. This has allowed me to assemble an artificial recreation of natural language with a lot of expressive power and flexibility.

Naturalistic programming languages have been attempted before with varying degrees of success and depth. My contribution to this field lies not in new goals but in the different strategies employed in my execution. Most projects I have found in this vein attempt to mimic and/or understand natural language in its entirety. This allows for a broader use of other natural languages or more idiomatic speech. \langName{} differs from these since it is first a programming language with the appearance of natural language secondary. The main benefit of being accessible is preserved while the pitfalls of ambiguity and the incredible scope of natural language are avoided. \langName{} does not understand what you say but is rather a machine that does what it is told to do.

This approach does not offer as much expressive freedom as a fully implemented natural programming language but is still a useful tool. When you progress away from the machine level you lose efficiency. This project does not lose as much efficiency as a comprehensive naturalistic language since it is still a programming language (albeit with some very English-looking syntax). This is not to say it rivals C, but rather that it does not suffer as much from the translation of English or some other language into code.

@@

Many tools have looked to leverage natural language to various purposes. The language COBOL attempted to simulate English with it is syntax \cite{Liu2005MetaforVS}. This meant defining control structures in a way that matched English framing as shown in \lstRef{cobolLoop}.

@@

\lstRef{cobolLoop} somewhat mimics the English framing used to show repeated actions. Other languages, including Python, also have syntax that can be somewhat read like English. Readability is important for people to quickly and easily understand code. However, some programming languages gone the extra mile.

@@

Another approach I discovered was Pegasus \cite{Knöll2006PegasusFS}. Pegasus uses many advanced components including short-term memory to store context, long-term memory to store semantic information, and a way of matching all the components of the supplied sentence to those memories. It is a very sophisticated system that gets closer than many to the natural language ideal at the cost of some performance issues.

Pegasus and Metafor are closer to natural language than other programming languages in terms of their understanding of the input text. However other attempts simply merely try to mimic the appearance of natural language without the computer truly understanding the underlying linguistic phenomena the way Pegasus or Metaphor might. There have been many other examples of natural language approaching languages (multiple of which are listed in Knöll \& Mezini). Some of these experiments had interesting ideas applicable to my project. A handful of them allow for fluffy constructions like “I would like …” or “Please …” These increase the flexibility. Also, the ability to store the inflections of various words and assign specific meanings to them is very interesting. Not all of these features were feasible to add, but some more fundamental pieces like alternate grammatical frames and a semblance of real semantic composition make \langName{} more like English than many standard programming languages.

\newpage{}
\section*{Goals}
The goals of \langName{} could be grouped into a few key groups: computational expressivity, natural language freedom, and linguistic fidelity. Balancing these three factors became a constant tug of war. Compromises had to be made in many places, but this push and pull of these three domains kept \langName{} focused on its true objective---becoming a programming language that looks like a natural language.

@@

\langName{} being Turing complete is as good as any programming language can get. Instead of proving that a language can simulate a Turing machine, one can also look at its features checking for branching control flow (if statements), looping control flow (while loops), and memory storage (vairables). If all these features are present, then the language is Turing complete.

@@

\subsection*{Linguistic Fidelity}
The last major goal was to incorporate as much real linguistic theory as possible to allow for increased language freedom and to offer a strong base for the parsing strategies \langName{} employs. This goal is actually really important when trying to work with the first two. To allow for language freedom, I could have implemented a large number of manually created frames for all sorts of operations, but this would not be scalable and would make the computational expressivity limited to whatever those frames did. My actual approach leverages the lambda-calculus nature of natural language semantics to allow for exceptionally expressive sentences that are very English-like.

\newpage{}
\section*{Planning}
Creating \langName{} has been a long and storied process with many ups and downs. The first step, as with many projects, is researching what others have done in the field. As mentioned earlier, the history of natural-looking languages is an eventful one. Starting with the advent of programming languages, programmers quickly began to bridge the computer-human communication gap. NLC is an unimplemented language designed by Bruce Ballard and Alan Biermann \cite{Ballard1979ProgrammingIN}. Their paper outlines a compelling argument for components a natural looking programming language needs and the types of data that such a language should be able to handle. Their model included a scanner, a parser, a semantic engine, and a matrix computer. \langName{} has the first three of these pieces as well. I utilize the ts-parsec library's built-in lexer and parser tools to tokenize the incoming sentences and then generate a pseudo-syntax tree. I can then run the top node of the tree and it will compose the meanings from across the syntax trees.

@@

Pegasus actually interpreting its input is in stark contrast with another popular avenue in this field A.I. Artificial intelligence, large language models (LLMs) in particular, has grown incredibly strong in recent years in terms of its ability to understand natural language prompts. There are many tools now from GitHub Copilot to Claude AI that leverage these models and large amounts of online data to generate code for you based on simple pseudo-code-esque prompts. I knew from the beginning however, that this is not the direction I wanted to go. A programming language by definition, should make the computer do what its told to do. You should be able to tell the computer exactly what you want it to do. This should be true relative to the level of abstraction inherent to your language. In C++, you have to worry about pointers and memory allocation, and whatever you tell it to do, it will do (even if it inevitably causes a seg. fault). In Python, you do not need to worry about memory management, but if you tell it to print something it will do exactly that.

I did not want to use LLMs because they don't do exactly what you tell them to do. One of these models will grab bits and pieces from all over the web to cobble together that only probably will do what you want, and even if it does work, there's no guarantee it will do it in an efficient, or secure manner \cite{Fu2023SecurityWO}. This is a big factor in my goal of computational expressivity. Any user should be able to both write what they want the machine to do and have it completed accurately, and another should be able to read that person's code and know what it will do. This drove me in the direction of natural language parsers.

@@

The last big step for beginning on the final version of \langName{} was some experiments with the context free grammar (CFG) parser from the NLTK library. I was successfully able to print strings and numbers as well as add and multiply numbers together.

@@

The last decision I had to make before creating the final implementation was which language to write in. I had started this project with Python to roughly throw some ideas at the wall. Getting deeper into the weeds led me to consider a statically typed language moving forwards. This would lead to more understandable code and less technical debt. For these reasons I decided to use TypeScript. This came with the added benefit of being able to use the React framework to create a streamlined and useful user interface.

\newpage{}
\section*{Product}
\subsection*{The Lexer}
The first layer of the \langName{} architecture is the lexer. This is a tool built using the ts-parsec library that groups the characters in the input into useful chunks. For my language, this meant words, numbers, and punctuation. Once I had these pieces I could further process them using the parser layer.

@@

The argument frame level is probably one of the most fundamentally important pieces of this entire project. This is the linchpin that allows for the natural language freedom that is so necessary in this project. \langName{} reads sentences by first capturing a verb symbol, and then repeatedly accepting arguments. In order to account for the basic verbs that I have included in the default lexicon for \langName{}, I thought of a number of ways in which the arguments can be supplied to the verb using the rest of the sentence. The big example of flexibility comes in the form of variables. A key metric I used when judging a good framing was whether or not it sounded like English. Originally I just used variable names like variable names in other languages. However, telling somebody to "Say stuff" in English would have a person either say the word "stuff" or they might just say random words. It felt unintuitive for any old word to be usable as a variable name when English words already have their own meanings. Thus the ways to access variable values in \langName{} are ``the value of \textlangle\textit{varName}\textrangle" and ``\textlangle\textit{varName}\textrangle's value." My new guiding principle was not just whether it sounded like English but whether somebody new to programming would inherently understand what was happening. Telling somebody to "Save 2 as bob's value" makes a lot more sense than "Save 2 as bob."

@@

\subsection*{The XBar Structure}
Now let us take a moment to explain how the XBars work in \langName{}. Every XBar node has a few key properties. First, it has a value. This value could be something like a string, a number, or boolean, or it could be something more complicated like a closure, or some more complicated function. That's where the second important piece of an XBar comes in. The type of an XBars value is stored using a custom Semantic Type object. These come in three varieties: the simple type (string, void, number, etc.), the variable type (contains a dictionary of string:Semantic Type pairs), the compound type (takes in a type and returns a type). By composing these types together, any type of an XBar can be described. The types can also be used to figure out whether nodes are able to compose with their sister nodes. Some types are also fuzzy which means they can match with multiple other types. The last important piece of data in an XBar is its children nodes. This set of other nodes are used to define its own value through semantic composition.

The XBar system I have employed is the primary way through which I adhere to the linguistic fidelity goal. By using semantic composition to derive the meaning of a sentence, these lines of code are actually working a lot more similarly to natural langauge than something like Python. Regular programming languages use CFG parsers to generate their code, but this is done through specific uses of parentheses, dots, and brackets. My strategy allows you to reuse other frames so long as they yield the correct types to compose.

@@

\subsection*{The Lexicon}
The lexicon is where the meanings of the sentence come from. The root of every sentence in \langName{} is the verb. This verbs definition comes from the lexicon tied to the scope the sentence is found within. Because the scope can change, everytime we reference a variable or verb, we need to make sure that the correct lexicon is being checked. Also, when a word was used, only the latest definition should be used. This meant that instead of creating an XBar with a number or string value whenever a variable name is referenced, instead we create an XBar that takes in a lexicon and returns a value. This was especially important considering loops. If you incremented  a value multiple times via a loop, you would only interpret the addition command once. If that does an operation based only on the initial state, than every operation will change the value to be exactly the same and nothing will be looped. If you resupply the current lexicon to the function each time, then the correct values are used.

@@

The shell's lexicon contains the verb definitions. A child lexicon is created within the shell lexicon for the outer for loop with a local variable stored in "character's value." Then in the scenario where that value is "e," the inner for loop creates an even deeper nested child lexicon within the current parent. For the second iteration of the outer loop, character's value will be set to "b," and the inner loop will create a child lexicon. It does not reuse the old one. This diagram reflects the current state during the third iteration of the outer loop.

At any given point, you have access to the current lexicon. Some functions like "For" will send you into a child lexicon and then have you discard it afterwards. When referencing a verb or variable, the code will find the lowest version of that symbol in the current scope hierarchy. In this example, "Say" is only found at the top level, but when "character" is referenced, it only needs to go up one level.

\newpage{}
\section*{Discussion}
\langName{} is a complicated network of different systems that leverage computer science and linguistic principles and it fulfills my three core goals. First, with the inclusion of while loops, if statements, and variables, \langName{} is Turing complete. Meaning it satisfies the computational expressivity requirement. The deeply involved parser structure allows for the natural language freedom and helps facilitate the XBar system which in turn maintains linguistic fidelity. So, all the goals are accomplished. What makes this project useful?

@@

It is also critical to note that this language has lots of room to expand. So many verbs and argument frames could be added to this language. What's more, it would not be entirely difficult. Each verb only requires you to creat its fundamental XBar (i.e. a value and a type). The level of specificity could also get more granular in future iterations. My parser tends to chunk certain words together (e.g. ``each \textlangle\textit{iterator}\textrangle\ in,'' ``the value of \textlangle\textit{variable}\textrangle''). Taking this approach to its extremes, you could imagine a lexicon where ``each,'' ``in,'' ``the,'' and ``of'' all have individual meanings. The other thing to note related to this is the fact that my parser for the most part only does branching in a certain direction. The verb is attached to the thing to its right, and then the next thing to the right is attached to that and so on. This means that complex argument structures don't really work. If you were to change how the parser worked to take into account other rules (e.g. phrase structure rules) or the ability for two words to semantically compose, you could get a lot closer to English with the programming language.

>>>>

% LANGUAGE NAME
\newcommand{\langName}{Vizzini}

@@

% MAIN BODY
\section*{Background}
The simplest way to describe my project is a programming language which appears naturalistic. The original programming languages were punch-card based and tied in to the machinery they worked on. Since then, new languages have evolved from low level machine code into higher level abstracted systems. The evolution from Fortran to C to Python (to name a few) showcases how languages have moved in a direction allowing for programmers to more abstractly tell the computer what to do. \langName{}\footnote{Named for the character in \textit{The Princess Bride} because of the language's verbosity and because sometimes things "[do not mean] what you think [they mean]"\cite{Goldman_1987}} is the natural next step in that evolution.

@@

While a fully implemented natural programming language has the potential to solve many problems, the development of \langName{} pimarily focused on the problem of accessibility. Software developers are the first thought when it comes to programming, but many others need to use computers to accomplish complicated tasks in research and other areas. Many researchers use various programs to collect and analyze data, run simulations, and perform other tasks. However, for many, programming is not familiar or easy to grasp. In my own experience teaching programming to children at a Summer camp, things as elementary as variables can be incredibly baffling when you've never been exposed to them.

@@

I have delved deep into parsers and semantic theory and learned how to make them work together. Having taken classes in syntax, semantics, and compilers, this is not an entirely new area for me, but my understanding of these topics has expanded beyond the scope of those classes. This has allowed me to assemble an artificial recreation of natural language with a lot of expressive power and flexibility.

Naturalistic programming languages have been attempted before with varying degrees of success and depth. My contribution to this field lies not in new goals but in the different strategies employed in my execution. Most projects I have found in this vein attempt to mimic and/or understand natural language in its entirety. This allows for a broader use of other natural languages or more idiomatic speech. \langName{} differs from these since it is first a programming language, the appearance of natural language being secondary. The main benefit of being accessible is preserved while the pitfalls of ambiguity and the incredible scope of natural language are avoided. \langName{} does not understand what you say but is rather a machine that does what it is told to do.

This approach does not offer as much expressive freedom as a fully implemented natural programming language but is still a useful tool. When you progress away from the machine level you lose efficiency. This project does not lose as much efficiency as a comprehensive naturalistic language since it is still a programming language (albeit with some very English-looking syntax). However, that is not to say it rivals C, but merely falls a bit closer to C than a language that suffers from the translation of English or some other language into code.

@@

Many tools have looked to leverage natural language to various purposes. The language COBOL attempted to simulate English with its syntax \cite{Liu2005MetaforVS}. This meant defining control structures in a way that matched English framing as shown in \lstRef{cobolLoop}.

@@

\lstRef{cobolLoop} somewhat mimics the English structure for talking about repeated actions. Other languages, including Python, also have syntax that can be somewhat read like English. Readability is important for people to quickly and easily understand code. However, some programming languages gone the extra mile.

@@

Another approach was Pegasus \cite{Knöll2006PegasusFS}. Pegasus uses many advanced components including short-term memory to store context, long-term memory to store semantic information, and a way of matching all the components of the supplied sentence to those memories. It is a very sophisticated system that gets closer than many to the natural language ideal at the cost of some performance issues.

Pegasus and Metafor are closer to natural language than other programming languages in terms of their understanding of the input text. However other attempts simply try to mimic the appearance of natural language without the computer truly understanding the underlying linguistic phenomena the way Pegasus or Metaphor might, and these are only a few of the natural language programming languages that have been created (multiple of which are listed in Knöll \& Mezini).

Some of these experiments had interesting ideas applicable to my project. For example, a handful of them allow for fluffy constructions like “I would like …” or “Please …” These increase language flexibility and fluidity. Also, the ability to store the inflections of various words and assign specific meanings to them is very interesting. Not all of these features were feasible to add, but some more fundamental pieces like alternate grammatical frames and a semblance of real semantic composition make \langName{} more like English than many standard programming languages.

\section*{Goals}
The goals of \langName{} could be grouped into a few key groups: computational expressivity, natural language freedom, and linguistic fidelity. Balancing these three factors became a constant tug of war. Compromises had to be made in many places, but this push and pull of these three domains kept \langName{} focused on its true objective---becoming a programming language that looks like a natural language.

@@

In terms of expressive power, \langName{} being Turing complete is as good as any language can get. To prove that a language can simulate a Turing machine, instead of actually programming one, you can look at its features checking for branching control flow (if statements), looping control flow (while loops), and memory storage (variables). If all these features are present, then the language is Turing complete.

@@

\subsection*{Linguistic Fidelity}
The last major goal was to incorporate as much real linguistic theory as possible to allow for increased language freedom and to offer a strong base for the parsing strategies \langName{} employs. This goal is critical when working with the first two. To allow for language freedom, I could have implemented a large number of manually created frames for all sorts of operations, but this would not be scalable and would make the computational expressivity limited to whatever those frames did. My actual approach leverages the lambda-calculus nature of natural language semantics to allow for exceptionally expressive sentences that are very English-like.

\section*{Planning}
Creating \langName{} has been a long and storied process with many ups and downs. The first step, as with many projects, is researching what others have done in the field. As mentioned earlier, the history of natural-looking languages is an eventful one. Starting with the advent of programming languages, programmers quickly began to bridge the computer-human communication gap. NLC is an unimplemented language designed by Bruce Ballard and Alan Biermann \cite{Ballard1979ProgrammingIN}. Their paper outlines a compelling argument for components a natural looking programming language needs and the types of data that such a language should be able to handle. Their model included a scanner, a parser, a semantic engine, and a matrix computer. \langName{} has the first three of these pieces as well. I utilize the ts-parsec library's built-in lexer and parser tools to tokenize the incoming sentences and then generate a pseudo-syntax tree. I can then run the top node of the tree and it will compose the meanings from across the syntax trees.

@@

Pegasus actually interpreting its input is in stark contrast with another popular avenue in this field AI. Artificial intelligences, large language models (LLMs) in particular, have grown incredibly strong in recent years in terms of their ability to understand natural language prompts. There are many tools now from GitHub Copilot to Claude AI that leverage these models and large amounts of online data to generate code for you based on simple pseudo-code-esque prompts. I knew from the beginning however, that this was not the direction I wanted to go. A programming language by definition, should make the computer do what its told to do. You should be able to clearly delineate the steps you want the machine to take. This should be true relative to the level of abstraction inherent to your language. In C++, you have to worry about pointers and memory allocation, and whatever you tell it to do, it will do (even if it inevitably causes a seg. fault). In Python, you do not need to worry about memory management, but if you tell it to print something it will do exactly that.

I did not want to use LLMs because they do not generate code that performs specific actions. Rather, these models will grab bits and pieces from all over the web to cobble together something that will only probably do what you want, and even if it does work, there's no guarantee it will do it in an efficient, or secure manner \cite{Fu2023SecurityWO}. This is a big factor in my goal of computational expressivity. Any user should be able to both write what they want the machine to do and have it completed accurately, and another should be able to read that person's code and know what it will do. This drove me in the direction of natural language parsers.

@@

The last big step before beginning on the final version of \langName{} was some experiments with the context free grammar (CFG) parser from the NLTK library. I was successfully able to print strings and numbers as well as add and multiply numbers together.

@@

The last decision I had to make before creating the final implementation was which language to write in. I had started this project with Python to roughly throw some ideas at the wall. Getting deeper into the weeds led me to consider a statically typed language moving forwards. This would help me to write more understandable code and generate less technical debt. For these reasons I decided to use TypeScript. This came with the added benefit of being able to use the React framework to create a streamlined and useful user interface.

\section*{Product}
\subsection*{The Lexer}
The first layer of the \langName{} architecture is the lexer. This is a tool built using the ts-parsec library that groups the characters in the input into useful chunks. For my language, this meant words, numbers, and punctuation. Once I had these pieces I could further process them using the parser layer.

@@

The argument frame level is probably one of the most fundamentally important pieces of this entire project. This is the linchpin that allows for the natural language freedom that this language is trying to achieve. \langName{} reads sentences by first capturing a verb symbol, and then repeatedly accepting arguments. In order to account for the basic verbs that I have included in the default lexicon for \langName{}, I thought of a number of ways in which the arguments can be supplied to the verb using the rest of the sentence. The big example of flexibility comes in the form of variables. A key metric I used when judging a good framing was whether or not it sounded like English. Originally I just used variable names like variable names in other languages. However, telling somebody to "Say stuff" in English would have a person either say the word "stuff" or they might just say random words. It felt unintuitive for any old word to be usable as a variable name when English words already have their own meanings. Thus the ways to access variable values in \langName{} are ``the value of \textlangle\textit{varName}\textrangle" and ``\textlangle\textit{varName}\textrangle's value." My new guiding principle was not just whether it sounded like English but whether somebody new to programming would inherently understand what was happening. Telling somebody to "Save 2 as bob's value" makes a lot more sense than "Save 2 as bob."

@@

\subsection*{The XBar Structure}
Now let us take a moment to explain how the XBars work in \langName{}. Every XBar node has a few key properties. First, it has a value. This value could be something like a string, a number, or boolean, or it could be something more complicated like a closure, or some more complicated function. That's where the second important piece of an XBar comes in. The type of an XBar's value is stored using a custom Semantic Type object. These come in three varieties: the simple type (string, void, number, etc.), the variable type (contains a dictionary of string:Semantic Type pairs), and the compound type (takes in a type and returns a type). By composing these types together, any type of an XBar can be described. The types can also be used to figure out whether nodes are able to compose with their sister nodes. Some types are also fuzzy which means they can match with multiple other types. The last important piece of data in an XBar is its children nodes. For a non-terminal node, semantic composition is used to generate its value based on these child nodes.

The XBar system I have employed is the primary way through which I adhere to the linguistic fidelity goal. By using semantic composition to derive the meaning of a sentence, these lines of code are actually working a lot more similarly to natural langauge than something like Python. Regular programming languages use CFG parsers to generate their code, but this is done through specific uses of parentheses, dots, and brackets. My strategy allows you to use any frames you want so long as they yield the correct types to compose.

@@

\subsection*{The Lexicon}
The lexicon is where the meanings of the sentence come from. The root of every sentence in \langName{} is the verb. This verbs definition comes from the lexicon tied to the scope the sentence is found within. Because the scope can change, everytime we reference a variable or verb, we need to make sure that the correct lexicon is being checked. Also, when a word is referenced, only the latest definition should be used. This meant that instead of creating an XBar with a number or string value whenever a variable name is referenced, instead we create an XBar that takes in a lexicon and returns a value. This was especially important considering loops. If you incremented  a value multiple times via a loop, you would only interpret the addition command once. If that does an operation based only on the initial state, than every operation will change the value to be exactly the same and nothing will be looped. If you resupply the current lexicon to the function each time, then the correct values are used.

@@

The shell's lexicon contains the verb definitions. A child lexicon is created within the shell lexicon for the outer for loop with a local variable stored as ``character's value.'' Then in the scenario where that value is ``e,'' the inner for loop creates an even deeper nested child lexicon within the current parent. For the second iteration of the outer loop, character's value will be set to ``b,'' and the inner loop will create a child lexicon. It does not reuse the old one. This diagram reflects the current state during the third iteration of the outer loop.

At any given point, you have access to the current lexicon. Some functions like ``For'' will send you into a child lexicon and then have you discard it afterwards. When referencing a verb or variable, the code will find the lowest version of that symbol in the current scope hierarchy. In this example, ``Say'' is only found at the top level, but when ``character'' is referenced, it only needs to go up one level.

\subsection*{Features}
\langName{} has a large number of features. These include:
\begin{itemize}
	\item Variable storing. Variables can be stored as ``the value of'' some name. In the future, it would be easy to extend this to objects by allowing more properties than just a ``value''. These variables can be stored and accessed using two different frames. First is the ``the value of [NAME]'' form and the second is the ``[NAME]'s value'' form.
	\item Printing. You can print values in \langName{} by telling it to ``Say'' something.
	\item If statements. There are actually two ways to use an if statement. The first is your classic if statement with an ``If'', a condition, and a colon followed by the if statement body. You can also do a terminal if statement, placing ``if'' and a condition at the end of your sentence.
	\item For loops. Like if statements, there is a classic for loop and a terminal for loop. Currently classic for loops are only used to iterate through a string. You start with ``For each'', then a name for the local iterator variable, then ``in'', and then the string you are iterating through followed by a colon and the loop body. The terminal for loop allows you to put ``[N] times'' at the end of your sentence in order to just run a simple sentence multiple times. `N' in this case could be a number or a variable with a number value.
	\item While loops. While loops have only a classic variant currently and follow the form ``While'', a condition, a colon, and then the loop body.
	\item Blocks. To create a connected block of code (e.g. to be used in a loop or if statement body), you only need to join the sentences together with ``, and'' instead of periods.
	\item Adding. The main operation that has been implemented in \langName{} is addition. You can add a number to a variable. Because negative numbers are valid, this means that subtraction is also implemented implicitly.
\end{itemize}

\section*{Discussion}
\langName{} is a complicated network of different systems that leverage computer science and linguistic principles and it fulfills my three core goals. First, with the inclusion of while loops, if statements, and variables, \langName{} is Turing complete. Meaning it satisfies the computational expressivity requirement. The deeply involved parser structure allows for the natural language freedom and helps facilitate the XBar system which in turn maintains linguistic fidelity. So, all the goals are accomplished. What makes this project useful?

@@

It is also critical to note that this language has lots of room to expand. So many verbs and argument frames could be added to this language. What's more, it would not be entirely difficult. Each verb only requires you to creat its fundamental XBar (i.e. a value and a type). The level of specificity could also get more granular in future iterations. My parser tends to chunk certain words together (e.g. ``each \textlangle\textit{iterator}\textrangle\ in,'' ``the value of \textlangle\textit{variable}\textrangle''), but taken to extremes, you could imagine a lexicon where ``each,'' ``in,'' ``the,'' and ``of'' all have individual meanings. The other thing to note related to this is the fact that my parser for the most part only does branching in a certain direction. The verb is attached to the thing to its right, and then the next thing to the right is attached to that and so on. This means that complex argument structures don't really work. If you were to change how the parser worked to take into account other rules (e.g. phrase structure rules) or use a parsing algorithm accounting for composability with backtracking, you could get a lot closer to English with the programming language.
